---
title: "Iris_PCA_test"
author: "Yisong Tao"
date: "11/18/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Iris PCA test

Use Iris dataset to check how much principle components change after removing
1.a noise variable; 2.a redundant variable; 3.a unique variale.

The purpose of this experiment is to see if we can use MCA to remove "bad" variables.

```{r setup dataset}
set.seed(0)
data_iris <- iris[,-5]   #load iris data, drop the factor variable
data_iris$noise <- rnorm(150, mean = 1, sd = 1) #add a noise variable
data_iris$PL2 <- 2*jitter(data_iris$Petal.Length) + 1 #add a redundant variable
data_iris$SW_PW <- 0.5*jitter(data_iris$Sepal.Width) + 0.5*jitter(data_iris$Petal.Width) #add another redundant variable
```

## PCA tests

PCA of all variables:
```{r pca}
library(psych)
prcp <- principal(data_iris, nfactors = 7, rotate = "none")
prcp$values
```

Remove a noise variable and run PCA
```{r pca2}
data_iris2 <- data_iris[,-5]
prcp_no_noise <- principal(data_iris2, nfactors = 6, rotate = "none")
prcp_no_noise$values
```
## Removing a varaible will reduce the sum of all principle components by 1, but if the variable is complete noise, the eigenvalue of important principle components won't change much.

Remove a redundant variable and run PCA
```{r pca3}
data_iris2 <- data_iris[, -3] # remove Petal.Length variable
prcp_noPL <- principal(data_iris2, nfactors = 6, rotate = "none")
prcp_noPL$values
```
```{r pca3_1}
data_iris2 <- data_iris[, -2] # remove Sepal.Width variable
prcp_noSW <- principal(data_iris2, nfactors = 6, rotate = "none")
prcp_noSW$values
```
```{r pca3_2}
data_iris2 <- data_iris[, -4] # remove Petal.Width variable
prcp_noPW <- principal(data_iris2, nfactors = 6, rotate = "none")
prcp_noPW$values
```

## Removing a redundant variable will not change the sum of eigenvalues of important principle components.


Remove a unique variable
```{r pca4}
data_iris_new <- cbind(data_iris, 0.01*c(1:150)) #adding a unique variable
prcp_new <- principal(data_iris_new, nfactors = 8, rotate = "none")
prcp_new$values
```
compare to:
```{r}
prcp$values
```
## Removing a unique variable will reduce the sum of eigenvalues of important principle components (by 1).

## We can use MCA to determine if we can drop a redundant or noise variable.
